---
title: "Lab 3: The 12 Days of Christmas"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

library(tidyverse)
```

## Setup

The song "12 Days of Christmas", written around 1780, tells the tale of many gifts a person recieves in the days leading up to Christmas. [https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)](https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)) These gifts repeat and compound; on the first day, the narrator recieves

```
A partridge in a pear tree.
```

On the last day, he or she recieves

```
Twelve Drummers Drumming
Eleven Pipers Piping
Ten Lords a Leaping
Nine Ladies Waiting
Eight Maids a Milking
Seven Swans a Swimming
Six Geese a Laying
Five Golden Rings
Four Calling Birds
Three French Hens
Two Turtle Doves
And a Partridge in a Pear Tree
```

Run the code below to load up a dataset called `xmas` that contains the crucial information about these gifts.  Your task will be to write a function that makes R sing the whole song to you, using only this dataset.

```{r}

xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

Throughout this lab, you are asked to write several functions.  You should write these functions in a `.R` script, NOT in a `.Rmd` markdown file.  When you want to test and use the functions, you should use `source("myscript.R")` to load the functions, as we did in PA 7.

If your functions depend on the tidyverse, or other packages, make sure to include appropriate `library` statements at the top of your `.R` script file.

Note that your functions can - and should! - include each other.  That is, don't duplicate code; reference earlier, smaller functions inside your larger functions.

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day 5 the narrator recieves "5 golden rings", but the entry in the dataset for the gift on day 5 simply says "ring".  Write a function called "pluralize_gift" that takes a gift and returns the appropriate plural. Some code structure is provided for you below.

*Hint 1:* Note that the gifts on days 6 and 9 have unusual pluralizations.

*Hint 2:*  The functions `paste()` and `paste0()` should be useful.  You will also likely want to use some `stringr` functions.*

*Hint 3:*  The following small example may be useful to you:

(`y$` is a special trick, called a regular expression, that says "match the letter y, but only if it is the last letter in the string")

```{r, eval = TRUE}
my_names <- c("Kimberly", "Trini", "Jason", "Billy", "Zach", "Tommy")

my_names %>% 
  str_replace("y$", "ies")
```


**Important:**  You should **absolutely not** "hard-code" anything into this function; this function should work in general, not just for the items in the 12 Days of Christmas.  For example, the word "rings" should not appear anywhere in the function.  I should be able to give it any gift and any day, and get back the plural.

```{r}

pluralize_gift <- function(gift){
  
  gift <- gift %>%
    str_replace(...)
  
  return(gift)
  
} 

```



Try your function out on the gifts in the dataset.  (Consider: is your function *vectorized*?)

```{r}

pluralize_gift(xmas$Gift.Item)

purrr::map_chr(xmas$Gift.Item, pluralize_gift)

```



## Step Two: Creating sentences

Write a function called `make_phrase` that takes as input the necessary information, and returns a phrase.  For example, 

`make_phrase(num = 10, item = "lords", verb = "a-leaping", adjective = "", location = "")`

should return

```
"10 lords a-leaping"
```

*Hint:*  The following may be useful to you:

```{r, eval = TRUE}
my_sentence <- "Oh no!     Too   many    spaces!    "

my_sentence %>%
  str_replace_all("\\s+", " ")

my_other_sentence <- c("1 bird", "11 birds")

my_other_sentence %>%
  str_replace_all("^1 ", "a ")

```

(What do you think the `^` symbol did?)

*Hint:* Look at the functions `str_replace_na()` and `str_trim` in the `stringr` package.


```{r}
make_phrase <- function(num, item, verb, adjective, location){
  
  
  
}


```


Try your function out on the `xmas` data, by making a new variable containing the daily phrases.

```{r}
xmas <- xmas %>%
  mutate(
    Full.Phrase = ...
  )
```



## Step Three:  Iteration.

Write a function called `sing_line` that takes as input:

* A dataset

* A number indicating which day to sing about

* The columns corresponding to the daily phrases.


```{r}
sing_line(xmas, 2, Full.Phrase)

```

should return

```
On the second day of Christmas, my true love sent to me,
two turtle doves 
and a partridge in a pear tree.
```

*Hint:* This function will need a loop, an `apply` function, or a `map` function inside of it, to iterate over all the gifts that are given on a particular day.

*Hint 2:*  The `{{phrase_col}}` part, which I provided for you, lets you use column names as arguments to a function.  Pretty neat!

```{r}
sing_line <- function(dataset, line, phrase_col){
  
  phrases <- dataset %>% pull({{phrase_col}})
   
  
  
  
}


```




## Step Four: Sing the song!

Use a function in the `map` or `apply` family to run the `sing_line` function over all 12 days of Xmas.

Note that, if done right, you are *repeating several lines* on each day.

Double check your output with the lyrics in the wikipedia link.


## Step Five:  One function to rule them all.

Combine steps 1-4 into one function called `sing_song`.  That is, this function should take as input only the dataset, and return the full song.

For this function - and only this function - you may assume that the names of the variables in your dataset are consistent.  That is, you may use the variable names `Gift.Item`, `Day.in.Words`, etc. in the body of your function.  However, the dataset name `xmas` should not appear in the body of your function.



Test your function on the following code, which should return the whole "12 Days of Christmas" song:

```{r, echo = TRUE}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")

sing_song(xmas)
```

## Flex Points

For up to 10 flex points, [document your functions thoroughly](https://style.tidyverse.org/documentation.html) using comments.


## Turn in your work

**READ THESE INSTRUCTIONS CAREFULLY!**

*There will be a -20% penalty for turning in the wrong files.*

1.  Download the file `Final_Song.Rmd` from Canvas

2.  Change the author name to your own name.

3.  In the second chunk, change the `"myscript.R"` part to successfully `source()` your own script with your singing functions.

4.  Knit `Final_Song.Rmd`

5.  Turn in, to Canvas:  (1) Your **script file** and (2) **Final_Song.html**.


