---
title: "Basics"
subtitle: "Day 2"  
author: "Stat 331"

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(flair)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

class: center, middle

# Functions

---
# Important R Facts

*Fact 1:* **R** is an **open-source** language.

--

*Fact 2:* The power of **R** comes from **packages**.

--

*Fact 3:* **R** is a **functional programming** language.

--

(This means it is designed for using and writing *functions*.)

---
# Anatomy of a function

```{r rnorm1}
rnorm(10, mean = 5, sd =3)
```

---
# Anatomy of a function

```{r rnorm, eval = FALSE}
rnorm(10, mean = 5, sd =3)
```

--
Function name:

```{r, echo = FALSE}
flair_chunk("rnorm", eval = FALSE) %>%
  hlt_quick("rnorm")
```

--

Required arguments:


```{r, echo = FALSE}
flair_chunk("rnorm", eval = FALSE) %>%
  hlt_quick("10")
```

(we often don't use the names for required arguments, but `n = 10` would be okay too)

--

Optional arguments:


```{r, echo = FALSE}
flair_chunk("rnorm", eval = FALSE) %>%
  hlt_quick("mean = 5") %>%
  hlt_quick("sd = 5")
```

(*always* use the argument names for optional arguments)

---
class: center, middle

# Troubleshooting Errors

---

# Syntax Errors

* Did you leave off a parenthesis?

```{r, echo = FALSE}
flair_code('rnorm(10, mean = 5', eval = FALSE) %>%
  hlt_fixed("(")
```

```{r, echo = FALSE, error = TRUE}

rnorm(10, mean = 5

```


---

# Syntax Errors

* Did you leave off a comma?

```{r, echo = FALSE, error = TRUE}
flair_code('rnorm(10 mean = 5)', eval = FALSE) %>%
  hlt_quick(" mean")
```

```{r, echo = FALSE, error = TRUE}

rnorm(10 mean = 5)

```


---

# Syntax Errors

* Did you make a typo?


```{r, echo = FALSE}
flair_code('rnrom(10, mean = 5)', eval = FALSE) %>%
  hlt_fixed("rnrom")
```

```{r, echo = FALSE, error = TRUE}

rnrom(10, mean = 5)

```

---
# Object type errors

* Are you using the right **input** that the function expects?

```{r, echo = FALSE}
flair_code("rnorm('hi', mean = 5)", eval = FALSE) %>%
  hlt_fixed("'hi'")
```

```{r, echo = FALSE, error = TRUE}

rnorm('hi', mean = 5)

```

---
# Object type errors

* Are you expecting the right **output** of the function?

```{r, echo = FALSE}
flair_code("my_obj <- rnorm(10, mean = 5) \n my_obj(5)", eval = FALSE) %>%
  hlt_fixed("my_obj(5)")
```

```{r, echo = FALSE, error = TRUE}

my_obj <- rnorm(10, mean = 5)

my_obj(5)

```

---
class: center, middle
# Object and Data Types

---

# Object Types

Is your object a...

* function?

* vector?

* list?

* data frame?

* matrix?

---
# Data Types

Does your data consist of...

--

* numbers?
    + integers?
    + doubles?
    
--

* words?
    + characters/string?
    + factor?
    
--

* something else???

---
class: center, middle
# (Fill in the grid activity)

---

class: center, middle

# R Markdown

---
# R Markdown

**R Markdown** is an *editor*.  (Like MS Word!)

It is built for combining **code chunks** with text.

--

Like any text document, you should **save** as you work!

--

When you are ready to see the finished product, you should **knit** your file.

---

class: center, middle

# Let's try it!
