---
title: "Functional Programming"
author: "Stat 331"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle, invert
# Iteration

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

```{r, include = FALSE}
library(tidyverse)
```
---

# Iteration

- What if you want to do something a whole bunch of times, or to every item in a list?  

```{r}
print("99 bottles of pop on the wall, 
      99 bottles of pop.  
      Take one down, pass it around, 
      98 bottles of pop on the wall.")

print("98 bottles of pop on the wall, 
      98 bottles of pop.  
      Take one down, pass it around, 
      97 bottles of pop on the wall.")

###.... and so on...
```

---
# We could write a function...

```{r}
sing_about_pop <- function(num) {
  
  
  paste(num,
    "bottles of pop on the wall,",
    num,
    "bottles of pop.",
    "Take one down, pass it around,",
    num - 1,
    "bottles of pop on the wall.")
  
}
```

---
# We could write a function...

```{r}
library(glue)

sing_about_pop <- function(num) {
  
  
  glue("{num} bottles of pop on the wall, {num} bottles of pop.
       Take one down, pass it around, {num - 1} bottles of pop on the wall.")
  
}
```

---
# This helped!

... but it is still repetitive...

```{r}
sing_about_pop(99)
sing_about_pop(98)
sing_about_pop(97)
```

---
# Good: For loops

```{r}
for (i in 99:95) {
  
  sing_about_pop(i)

}
```
--

... wait, where did it go?

---
# Print!

```{r}
for (i in 99:95) {
  
  print(sing_about_pop(i))

}
```

---
# Good: For loops

**Pro:**

* Classic programming approach; easy to understand

* Can change values at each loop step.


**Con:**

* Slow in R.

* A little messy to read the code

* Not always necessary.


---
# Better: The `apply` functions

```{r}

sapply(99:95, sing_about_pop)

```
---
# Better: The `apply` functions

- Pay attention to what type of input they *take* and what type of output they *give*

- `sapply` takes a vector, returns a vector

- `lapply` takes a list, returns a list

- `apply` takes a dataset and a dimension (1 = row, 2 = col), returns a vector or list

---
# Best:  The `map` functions from `purrr`


```{r}

map_chr(99:95, sing_about_pop)

```

---
# Best:  The `map` functions from `purrr`

- `map` always returns a **list**

- `map_chr` always returns a **character vector**  (or an error)

- What do you think `map_dbl` returns?

---

# What if I want some options in my function?

```{r}
sing_about_drinks <- function(num, drink) {
  
  
    
  glue("{num} bottles of {drink} on the wall, {num} bottles of {drink}.
       Take one down, pass it around, {num - 1} bottles of {drink} on the wall.")
  
}
```

---
# `sapply` has trouble!

```{r, error = TRUE}

sapply(99:95, sing_about_drinks(drink = "beer"))

```
---

# You can kind of get around this...

(This is called a **headless function**)

```{r}

sapply(99:95, function(x) sing_about_drinks(x, drink = "beer"))

```

---

# But `map` has a cool shortcut!

- Note the "~" in front of the function - this is important!

- Note the "." before the "x" - this is important!

```{r}

map_chr(99:95, ~sing_about_drinks(.x, "beer"))

```


---

# What if I want to use many options?

```{r, echo = FALSE}

data <- data.frame(num = 99:95, drink = c("pop", "beer", "water", "lemonade","milk"))

```

```{r}
head(data)
```

---
# Use `map2`!

```{r}
map2_chr(data$num, data$drink, ~sing_about_drinks(.x, .y))
```

---
# The very best: Vectorized functions

## What does it mean to be vectorized?

```{r}
add_odds <- function(x,y){
  
  if(x %%2 == 1 & y %% 2 == 1){
    
    x + y
    
  } else {
    
   0
    
  }
  
}

```

---

# What does it mean to be vectorized?

```{r}

add_odds(1,3)

add_odds(2,4)


```

---

# What does it mean to be vectorized?

```{r, error = TRUE }

first_nums <- c(1,2)
next_nums <- c(3,4)

add_odds(first_nums, next_nums)

```

---

# What does it mean to be vectorized?

`add_odds` is not a vectorized function!

```{r, eval = FALSE}
add_odds <- function(x,y){
  
  if(x %%2 == 1 & y %% 2 == 1){
    
    x + y
    
  } else {
    
   0
    
  }
  
}

```

---

# We can try to make it vectorized...
 
```{r}
add_odds_2 <- function(x,y){
  
  is_odd <- x %% 2 == 1 & y %% 2 == 1
  
  answers <- rep(0, length(x))
  
  answers <- x + y
  
  answers[!is_odd] <- 0
  
}

```

---
# We can try to make it vectorized...

```{r}

first_nums <- c(1,2)
next_nums <- c(3,4)

add_odds_2(first_nums, next_nums)

```

---
# But, what if we can't vectorize?
## Iteration!

```{r}

first_nums <- c(1,2)
next_nums <- c(3,4)

map2(first_nums, next_nums, ~add_odds(.x, .y))

```

---

## Things to remember

- What does your function take as INPUT?

- What does your function give as OUTPUT?

- How can we write it to be friendly with `apply` or `map`?

- Do you REALLY need to use `apply` or `map`, or is it vectorized?

- Bonus: How can we write it to be friendly with `%>%`?





