---
title: "PA 11:  Functions and Strings Review"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```

**This PA is optional!**

This is not the same as Extra Credit.  This means that if you complete it, it will be part of your overall PA grade average; but if you choose not to complete it, you will not be penalized.

## The Caesar Cipher

The Caesar Cipher is one of the earliest known methods of encrypting text.  The way it works is that every letter in the text is "shifted" a certain amount.  For example, a Caeasar cipher of `abc` with a shift of `4` would result in `efg`.

This activity will ask you to decode a secret message using the Caesar cipher.  You may want to start by googling around, just to get ideas about how this cipher works.

## Task One:  Function writing

Write a function called `caesar` that translates messages.  This function should take as arguments a single word (as a string) and a number to shift by (can be positive or negative).  It should return the word that has been shifted.

Hints:

* You will want to think about how your code should handle numbers that are negative/zero, and numbers that are larger than 26.
* There is a pre-loaded vector called `letters` that contains a-z.

If your function is successful, the code below should return a real word.

```{r}

library(tidyverse)

shift_letter <- function(letter, num){
  
  this_index <- which(letters == letter)
  new_index <- this_index + num
  
  new_index = new_index %% 26
  
  if(new_index == 0){
    new_index = 26
  }
  
  letters[new_index]
  
}

caesar <- function(word, num){
  
  letters_in_word <- unlist(str_split(word, ""))
  
  paste0(map_chr(letters_in_word, ~shift_letter(.x, num)), collapse = "")
  
}
```

```{r, echo = TRUE}

caesar("sfzqlov", -23)

```

## Task Two: Decode a message

The encrypted message can be loaded using the code below.  This message has been encrypted via the Caesar cipher, with an unknown shift.

```{r, echo = TRUE}

message <- read.csv("https://www.dropbox.com/s/ctplwmkvey4yo0e/cs_encrypted.txt?dl=1",
                    stringsAsFactors = FALSE, header = TRUE)$x
```


1. Use your `caesar` function on the first word only, trying all the possible shifts from 0 to 26.  Determine which shift was used for the full message.

```{r}
for(i in 1:26){
  
  caesar(message[1], i)
  
}
```


2. Use `map` or similar to decode the full message.

```{r}
translated_words <- map_chr(message, ~caesar(.x, 9))
str_c(translated_words, collapse = " ")
```


3. On Canvas, submit the name of the historical figure who gave the speech in the message.