---
title: "PA 9: Simulation"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
set.seed(1957)
```

For this assignment, set your seed to 1957.

A man named Professor Harold Hill wanders into your small town.  He claims to be selling top-quality instruments for a marching band.  He begins selling his instruments to all the children in town.

You are suspicious of Professor Hill, and think that perhaps he is selling instruments made of subpar materials, to scam the townsfolk.

You do some research on the weights of properly crafted brass instruments, and you learn the following facts:

* Trombones weights are Normally distributed, with a mean weight of 4.6 pounds and a standard deviation of 0.8 pounds.

* Trumpet weights are Uniformly distributed between 1.5 and 3.5 pounds.

* Reed instruments (like clarinets, bassoons, and saxophones) have weights are Exponentially distributed with a mean of 4 pounds.


## Warmup

* What is the 95th percentile for trumpet weight?

* What is the 10th percentile for trombone weight?

* About what percent of trombones do you expect to be more than 5 pounds?

* About what percent of reed instruments do you expect to be more than 5 pounds?

* Simulate 100 random trombone weights.  How many of them were below 4 pounds?

## Catching a con

You manage to intercept a shipping notice for delivery to Professor Hill.  It says the following:


```
Wells Fargo Shipment 1957

To:  Harold Hill, Mason City, Iowa
From:  Music Factory, Gary, Indiana

Items included:
    Trombones: 76
    Cornets: 110
    Saxophones: 542
    Clarinets: 318
    Bassoons: 175
    
Total Shipped Weight: 4532 lbs.

```

This is your chance to catch Professor Hill in his lie!

First, write a function that samples the correct number of weights of trombones, cornets (trumpets), and reed instruments, and then returns the total weight.

Then, use that function to create 1000 random samples of total weight.

How many of these samples had a weight less than Prof. Hill's shipment?

```{r}
music_man <- function(){
  
  trom <- rnorm(76, mean = 4.6, sd = 0.8)
  cor <- runif(110, 1.5, 3.5)
  reed <- rexp(1035, 1/4)
  
  return(sum(trom, cor, reed))
  
}

my_weights <- replicate(1000, music_man())
sum(my_weights < 4532)
```



