---
title: "Practice Activity 2: Objects and subsetting"
author: "YOUR NAME HERE"
output: html_document
---

### Part One: Data import and cleaning

First, we declare our package dependencies and load the data.

```{r setup, include = FALSE}
library(tidyverse)



colleges <- read.csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```



There are too many variables in this dataset.  We don't need all of them. Narrow your dataset down to only:

Name of the institution
City, State, and ZIP code of the institution
The Admissions Rate
The average SAT score
The number of undergraduate students
The in and out of state tuitions
Whether the school is public or private
The "REGION" variable.

```{r}
colleges_clean <- colleges %>
  select(INSTNM, CITY, STABBR, ZIP, CONTROL, ADM_RATE, SAT_AVG, TUITIONFEE_IN, TUITIONFEE_OUT, UGDS, REGION) 
```

Drop the schools that are private and for-profit (category 3).

```{r}
colleges_clean <- colleges_clean %>%
  filter(CONTROL == 1, CONTROL == 2) 
```

Adjust the appropriate variables to be numeric.

```{r}
colleges_clean <- colleges_clean %>%
  mutate(
    TUITIONFEE_IN = numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
    SAT_AVG = numeric(SAT_AVG),
    ADM_RATE = numeric(ADM_RATE)
    ) 
```

Adjust the appropriate variables to be factors.

```{r}
colleges_clean %>% mutate(
    CONTROL = as.factor(CONTROL),
    REGION = as.factor(REGION)
)
```

Create a new variable called TUITION_DIFF which contains the difference between in and out of state costs.

```{r}
colleges_clean <- colleges_clean %>% 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```


Drop all the rows with missing data.  
(This is not always a great idea!  Usually, even if *some* of the information is missing, we don't want to throw out the entire row.  This time, however, we'll be lazy.)

```{r}
colleges_clean <- colleges_clean %>% drop.na()
```



Lastly, notice that each of these steps started with 

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% ...
```

That is pretty redundant!  Instead, we could perform all these tasks as one long "pipeline".  Combine your (fixed) code chunks into a single chunk that cleans the data.




## Part Two:  Identify the mystery college

Wow!  Your best friend Ephelia has been accepted to the college of her dreams!  Unfortunately, Ephelia is a very mysterious person, and she won't tell you directly which college this is.  You'll have to use her clues to figure out which school is her dream school.


Clues:

1. This college is located in Region 1.

2. This college's admission rate is in the first quartile for the region.

3. This college charges the same for in- and out-of-state tuition.

4. The average SAT score of this college is an odd number.

5. This college is NOT in New Hampshire or in the city of Boston.

6. More than 3,000 people apply to this college every year.  (Assume the size of the freshman class is 1/4 of the undergraduate population.)

7. Dr. Bodwin did not attend this college.

8. Of the two options remaining at this step, Ephelia will attend the cheaper one.


