---
title: "PA 12:  Distributions and Bootstrapping Review"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
library(tidyverse)
```

**This PA is optional!**

This is not the same as Extra Credit.  This means that if you complete it, it will be part of your overall PA grade average; but if you choose not to complete it, you will not be penalized.

## The Cauchy Distribution

The Cauchy distribution is a continuous distribution with possible values from negative infinity to infinity.  It is an unusual distribution in many ways.  We will use simulation and bootstrapping to explore this.

## Instructions

Perform the following steps:

1. Simulate a sample of size 100 from the standard Cauchy distribution.  *(Hint: Remember the `p, q, r, d` functions for simulation!)*

2. Take 1000 bootstrap samples from your simulated sample.  Make a 95% confidence interval for (a) the mean and (b) the median.

```{r}

c_sam <- rcauchy(100)

csam <- replicate(1000, sample(c_sam, replace = TRUE), simplify = FALSE)

mans <- map_dbl(csam, mean)
meds <- map_dbl(csam, median)


quantile(mans, c(.025, .975))
quantile(meds, c(.025, .975))


```

3. Repeat (1) and (2) ten times.  That is, you should have a NEW simulated Cauchy sample each time.  Save the upper and lower bounds of the mean and median confidence intervals at each step.

Your result should be a data frame that looks something like this, but with very different values:

```{r, eval = TRUE}
fake_df <- tibble(
  trial = 1:5,
  upper_mean = runif(5),
  lower_mean = upper_mean - runif(5),
  upper_median = runif(5),
  lower_median = upper_median - runif(5),
)

fake_df
```


4. Make a plot showing the 10 mean confidence intervals, and another plot showing the 10 median confidence intervals.  (It is your choice what kind of plot to use - `geom_errorbarh` is a nice option.)

```{r}


my_CIs <- data.frame()

for(i in 1:10){
  
  c_sam <- rcauchy(100)
  
  csam <- replicate(1000, sample(c_sam, replace = TRUE), simplify = FALSE)
  
  mans <- map_dbl(csam, mean)
  meds <- map_dbl(csam, median)
  
  
  my_CIs <- rbind(my_CIs, c(quantile(mans, c(.025, .975)), quantile(meds, c(.025, .975))))

}

names(my_CIs) = c("lower_mean", "upper_mean", "lower_med", "upper_med")

ggplot(my_CIs) + geom_errorbarh(aes(xmin = lower_mean, xmax = upper_mean, y = 1:10))

ggplot(my_CIs) + geom_errorbarh(aes(xmin = lower_med, xmax = upper_med, y = 1:10))

```


## Conclusion

Mathematically, for the Cauchy distribution, one of the statistics we studied (mean or median) does not exist!  That is, it is equal to infinity, and therefore impossible to estimate with bootstrapping.

For one of your statistics, your bootstrap CIs should have been fairly similar.
For the other, they should be all over the place.

Use your results to answer the question on Canvas:  which statistic (mean or median) does not exist for the Cauchy distribution?
