---
title: "Practice Activity 2: Objects and subsetting"
author: "YOUR NAME HERE"
date: "January 13, 2019"
output: html_document
---

Today you will be creating and manipulating vectors, lists, and data frames to uncover a top secret message.

## Part One: Dealing with errors

This section will create a few objects for you to use in Part Two.  

Alas, each of the following R chunks will cause an error and/or do the desired task incorrectly.  You will need to find the mistake, and correct it, to complete the intended action. 

### Advice for dealing with errors

**Errors can be sneaky - check results often** 

If a chunk of code runs smoothly without giving you any error or warnings this does not necessarily mean it accomplished the desired task.

It is a good habit to check the results of your code every time you finish a task.

* In the text before the code chunk, make sure to briefly state what the point of the chunk is.  This will remind later readers - which might be your future self! - what the desired output is.

* If you created a new object, take a look at it, either by clicking its name in your *Environment* tab or by typing its name into the console.  Make sure it looks about how you expect.

* If you created or updated a data frame, make sure your edits did what you hoped.  Use the *Environment* or the `head()` function to investigate your changes.


**Two heads are better than one**

It can be hard to spot bugs in code that you wrote.  

Work with those around you - if something goes wrong, ask a friend to take a peek at your code and see if any glaring errors (like *syntax error*) pop out.

**Explain your code out loud**

The best way to troubleshoot a sneaky but is to explain out loud each step of your code, and what you hoped to accomplish.

If you are alone, try [Rubber Duck Debugging](https://en.wikipedia.org/wiki/Rubber_duck_debugging)!

**Google is your friend**

The whole of the internet is at your disposal!  Use it early, use it often.

Some tricks:

* Copy-paste the exact error message into Google. Chances are, somebody else had a similar problem and got a similar message.

* Include package names in your search terms.  For example, "bar plot in ggplot" is a better search than "bar plot in R".


### Fix these steps

1. Create vectors containing the upper case letters, lower case letters, and some punctuation marks.

```{r}
lower_case <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y", "z")

upper_case <- c("A", "B", "C", "D", "E", "F", "G", "H" "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z")

punctuation <- c(".", ",", "!", "?", "'", """, "(", ")", " ", "-", ";", ":")

```

2. Make one long vector containing all the symbols.

```{r}

my_symbols <- cbind(lower_case, upper_case, punctuation)

```

3. Turn the `my_symbols` vector into a data frame, with the variable name `Symbol`

```{r}
my_symbols <- dataframe(my_symbols)
names(my_symbols) <- Symbol
```


4. Find the total number of symbols we have in our data frame.  

```{r}
len <- length(my_symbols)
```


5. Create a new variable in your dataframe that assigns a number to each symbol.

```{r}
my_symbols%Num <- 1:len
```

## Part Two: Decoding the secret message.

You should now be read to decode the message.  Run this chunk to load up the encoded secret message as a vector:

```{r}

top_secret <- read.csv("https://raw.githubusercontent.com/kbodwin/STAT-331/master/In-Class%20Activities/2_Secret_Code?token=AVHCwR08vI1TB2BJE1euD15U1SD18E6pks5cP7F7wA%3D%3D", header = FALSE)$V1

```

By altering this top secret set of numbers, you will be able to create a message.  Write your own code to complete the steps below.  Each step should be a separate code chunk.  

At each step, you should be updating the vector `top_secret`, rather than creating many new vectors.  If you make a mistake, simply reload the data and run each chunk in order to catch back up.

Make sure you check in at each *Checkpoint* step as well, to see if you are on track.

1. Add 14 to every number.
2. Multiply every number by 18, then subtract 257.
3. Exponentiate every number.  (That is, do e^[number].  You may have to Google this!)
4. Square every number.

**Checkpoint:** Headquarters has informed you that at this stage of decoding, there should be 352 numbers in the secret message that are below 17.

5. Turn your vector of numbers into a matrix with 5 columns.
6. Separately from your top secret numbers, create a vector of all the even numbers between 1 and 382. Name it "evens". That is, "evens" should contain 2, 4, 6, 8 ..., 382.
7. Subtract the "evens" vector from the first column of your secret message matrix.
8. Subtract 100 from all numbers 18-24th rows of the 3rd column.
9. Multiply all numbers in the 4th and 5th column by 2.
10. Turn your matrix back into a vector.

**Checkpoint:** Headquarters has informed you that at this stage of decoding, all numbers in indices 500 and beyond are below 100.


11. Take the square root of all numbers in indices 38 to 465.
12. Round all numbers to the nearest whole number.
13. Replace all instances of the number 39 with 20.

**Checkpoint:** Headquarters has informed you that your final message should have 344 even numbers.


## Part 3: The secret message!

Run the code chunk below to decode your message.

```{r}
paste(my_symbols[top_secret], collapse = ""))
```


Google the first line of this message, if you do not recognize it, to see what it is. Submit your answer to the PA 2 assignment on Canvas.

