---
title: "Practice Activity: Loops, Functions, Conditionals"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This activity will ask you to create a script file with several functions in R. Apart from the quick check-ins, you will NOT need to edit any of the code in this .Rmd file - if your script is correct, it will run smoothly.

First, save this file somewhere in your project folder.  Then create an **R Script** (not R Markdown!) file called "PA_Script.R" to the main part of your project folder.  

Then follow the steps below, all in the file "PA_Script.R"

1. Write a function called `divide_and_round()`, which takes a number or vector of numbers, divides them all by their smallest value, and rounds the results to the nearest whole number.  Here is some starter code to work from:

```{r}
divide_and_round <- function(nums){
  
  my_mean <-1
  divided <- nums/my_mean
  rounded <- 0
  
  rounded
  
}
```


Test it by running the code below, which should return `1  2  2  2 NA`

```{r}
source("PA_Script.R")

test <- c(2:5, NA)

divide_and_round(test)
```


2. Write a function called `no_nines_or_twelves()`, which takes a number or vector of numbers and returns `TRUE` if the number is NOT divisible by 9 or 12, `FALSE` otherwise.


Test it by running the code below, which should return `TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE FALSE  TRUE    NA`

```{r}
source("PA_Script.R")

test <- c(1:10*5, NA)

no_nines_or_twelves(test)
```

3. Write a function called `every_other()` which takes a vector and returns every other value in the vector.  

Include an optional argument called "start" which lets you choose where to start skipping; that is, if `start = 1`, we return the 1st value, 3rd, 5th, etc. and if `start = 2`, we return the 2nd, 4th, 6th, etc. 

Some starter code for you:

```{r}
every_other <- function(vec, start = 1){
  
  good_indices <- ???
  
  vec[good_indices]
  
}
```


Double check your `every_other()` function, as we have done for the other functions above. Make sure the following returns what you think it should.

```{r}
source("PA_Script.R")

test <- c(1:10)

every_other(test)
every_other(test, start = 2)
```



4. Write a function called `shorten()` which takes a vector and keeps dropping the first value, until the sum of the vector is less than 350  Return the remaining values.

The code below should return: `150 100 50`


```{r}
source("PA_Script.R")

test <- rev(c(1:10*50))

shorten(test)
```



5. Once you have written your four functions correctly, run the following code:

```{r}

source("PA_Script.R")

my_vec <- c(39, 1.87, 48, 11, 8, 45, 21, 5, 12, 33, 9, 11, 108, 4, 18, 5, 16, 17, 8, 48, 27, 24, 4, 22, 12, 27, 23, 46, 42, 35, 15, 34, 36, 26, 18, 10, 18.21, 72.04, 36.9, 41.81, 29, 89.75, 34.03, 20.18, 48.74, 15.76, 31.86, 83.6, 43.55, 39.99, 23.55, 8.54, 24.71, 22.02, 9.71, 62.14, 35.46, 16.61, 15.66, 21.29, 30.52, 201.07, 45.81, 7.85, 30.13, 34.14, 22.62, 10.2, 6.02, 30.12, 10.64, 31.72, 24.57, 14.43, 43.37, 89.93, 44.72, 51.32, 13.62, 45.56, 22.96, 7.05, 29.99, 41.38, 26.59, 23.04, 19.82, 50.73, 39.56, 43.79, 30.22, 85.85, 5.78, 78.85, 29.52, 66.27, 44.06, 27.28, 24.43, 64.32, 3.35, 67.45, 46.72, 48.44, 48.65, 33.3, 40.28, 19.04)

my_vec <- every_other(my_vec, start = 2)
my_vec <- divide_and_round(my_vec)
my_vec <- every_other(my_vec, start = 1)
my_vec <- shorten(my_vec)
my_vec <- my_vec[no_nines_or_twelves(my_vec)]

my_vec <- sort(my_vec)

my_vec

```

If you have done everything correctly, your final vector will be six numbers long.  Google these six numbers to find the TV show reference.  Write the name of this show in your assignment on Canvas to get full credit for this activity.











