---
title: "Lab 5: Simulation"
author: "Stat 331"
output: rmdformats::readthedown
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=6, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)

library(tidyverse)

set.seed(369)
```


# The Central Limit Theorem

Recall that the **Central Limit Theorem** tells us about the distribution of a sample mean, which is called the sampling distribution.  As sample size becomes large, the sampling distribution approaches a normal distribution, and the standard deviation of the sampling distribution decreases. 

Let's take a look at what exactly all this means.

**For this assignment, use a seed of 369.**

## Part One:  The CLT

1.  Suppose we have a quantity that we know is Normally distributed with mean 50 and standard deviation 10. Below, I have generated 1000 samples from this variable, and plotted the density.

```{r, echo = TRUE}

my_samples <- data.frame(
  values = rnorm(1000, mean = 50, sd = 10)
)

ggplot(my_samples, aes(x = values)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) + 
  geom_vline(xintercept = mean(my_samples$values), color = "red", lwd = 2)

```

  +  What were the mean and standard deviation of the samples?
  
```{r}
my_samples %>%
  summarize(mean(values), sd(values))
```
  

  +  Simulate a sample of size 30 from this population, and find the mean and standard deviation.  How do they compare to your answers above?
  
```{r}
my_samples <- data.frame(
  values = rnorm(30, mean = 50, sd = 10)
)

my_samples %>%
  summarize(mean(values), sd(values))
```

```
The mean and sd should be close to the same for both answers.
```

```
The following is NOT correct:
```
```{r}
sample(my_samples$values, 30)
```



2.  Repeat Question 1, this time for a variable that has an `exponential(0.02)` distribution.

```{r}
my_samples <- data.frame(
  values = rexp(1000, rate = 0.02)
)

ggplot(my_samples, aes(x = values)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) + 
  geom_vline(xintercept = mean(my_samples$values), color = "red", lwd = 2)


my_samples %>%
  summarize(mean(values), sd(values))

my_samples_30 <- data.frame(
  values = rexp(30, rate = 0.02)
)

my_samples_30 %>%
  summarize(mean(values), sd(values))

```

```
The mean and sd should be close to the same for both answers.
```


3.  Still using the `exponential(0.02)`, simulate 1000 samples of size 30, and calculate the sample mean of each one.  Make a density plot of these *sample means*.

```{r}
many_samples <- replicate(1000, rexp(30, rate = 0.02), simplify = FALSE)

my_sample_stats <- data.frame(
  sample_means = map_dbl(many_samples, mean)
)

my_sample_stats %>%
  ggplot(aes(x = sample_means)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) 
  
```


4.  Comment on your plot from (3).  How was the distribution of the sample means different from the distribution of the original variable?

```
Roughly normally distributed, not as skewed.
```

5. Find the standard deviation of your 1000 sample means.  How does it compare to the standard deviation of your single sample from 2(b)? *Hint: Divide the sd of the single sample by the sd of the sample means.  Then square that.*

```{r}
sd(my_samples$values)
sd(my_sample_stats$sample_means)

## (exact values will vary)

(50/9.13)^2
```

```
The variance of the sample means is the variance of the single samples divided by n.
```

## Part Two: Beyond the Mean

The Central Limit Theorem works "like magic" when dealing with sample means.  But what about other statistics?

Continue to study the `exponential(0.02)` distribution.

1. The Median

  + What is the true median of this distribution?
  
```{r}
qexp(.5, 0.02)
```

```
NOT CORRECT:
```

```{r}
median(my_samples$values)
```
  

  + Return to your 1000 samples of size 30.  Find the median of each sample.  Make a plot of the density of all the medians, with a vertical line representing the true median.
  
```{r}
my_sample_stats <- my_sample_stats %>% 
  mutate(
    sample_medians = map_dbl(many_samples, median)
  )

my_sample_stats %>%
  ggplot(aes(x = sample_medians)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) + 
  geom_vline(xintercept = qexp(.5, 0.02), color = "red", lwd = 2)
```
  

  + Based on your plot above, do you think the CLT applies to the median of an exponential distribution?
  
```
Comment on whether or not it looks Normally distributed around a center at the true value of the median.
```
  

2. Repeat Q1, this time for the *midhinge* of each sample.  (You may have to google this!)

```{r}
mh <- (qexp(.75, 0.02) + qexp(.25, 0.02))/2
mh
```

  
```{r}
midhinge <- function(sample) {
  
  q1 <- quantile(sample, .25)
  q3 <- quantile(sample, .75)
  
  return((q1+q3)/2)
  
}


my_sample_stats <- my_sample_stats %>% 
  mutate(
    sample_midhinges = map_dbl(many_samples, midhinge)
  )

my_sample_stats %>%
  ggplot(aes(x = sample_midhinges)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) + 
  geom_vline(xintercept = mh, color = "red", lwd = 2)
```
  
  
```
Comment on whether or not it looks Normally distributed around a center at the true value of the midhinge.
```

3. Repeat Q1, this time taking the *maximum* of each sample.  (There is no coding way to find the true maximum; make your best guess.)

```{r}
qexp(1, 0.02)
```

```
There is no maximum!
```
  
```{r}
my_sample_stats <- my_sample_stats %>% 
  mutate(
    sample_maxes = map_dbl(many_samples, max)
  )

my_sample_stats %>%
  ggplot(aes(x = sample_maxes)) + 
  geom_histogram(aes(y = ..density..),
                 fill = "white",
                 color = "black") + 
  geom_density(fill = "cornflowerblue", alpha = 0.5) 
```
  
  
```
No!  Even if it looks a little bell-shaped, it can't possibly be centered at the true value (infinity).
```


## Part Three:  Appreciate the usefulness

You have been told that the amount of time you have to wait for a bus from Cal Poly to Downtown SLO is `exponential(0.02)`.

You think this might be a lie.  In the last 30 days, you have waited for the bus an average of 55 minutes.

If the bus system is telling the truth about the `exponential(0.02)` distribution, how unlucky would you have to have been this month?

Make an argument both by direct simulation, and by using the Central Limit Theorem.

(*Hint: The mean of `exponential(0.02)` is `50`, and the sd is equal to the mean.*)

#### Using simulation:

We've already simulated 1000 samples of size 30 from this population!

```{r}
sum(my_sample_stats$sample_means > 55)/1000
```


#### Using CLT:

The sample mean is approximately Normal, with a mean of 50 and a sd of 50/sqrt(30).

```{r}
pnorm(55, mean = 50, sd = 50/sqrt(30))
```


## Flex Points:

Use Google or other resources to learn about the **Normal approximation to the Binomial distribution**.

Write a very brief explanation of this phenomenon, that makes use of simulation and visualization to illustrate the concept.

Upload your writeup **separately** to Canvas.
